### 2016-02-07 prepReport/theorBack.tex
Corrected some mistakes in the report

\section{Basic Concepts of Machine Learning}
Machine Learning is a subject of Computer Science that studies the development of mathematical formulations that allows software to learn autonomously, without an explicit description of each rule of operation. Its goal is to extract latent characteristic features in datasets that allow an immediate classification of each input data into a particular class -- the catch is that there is no previous rule formulation, but instead we have an adaptive model that adjusts is parameters automatically from the input data it receives.

Let us consider a practical example. For instance, suppose we want to build a device that given an input audio waveform representation of a spoken word, it matches it into a particular word of a dictionary. We could, of course, devise a set of rules and exceptions for each word analysing some of its features (perhaps the Fourier representation of each one, and, from it, manually finding the appropriate rules for each), but apart from it being a very complex task, it wouldn't be a scalable solution, given the enormous number of words in each language. The approach taken by Machine Learning is diametrically different, and instead of manually processing each waveform, we build a large dataset, of size $N$, containing the waveforms of several words $\left[ \mb{x}_1 \; \mb{x}_2 \;\mb{x}_2 \; \dots \mb{x}_N \right]$ -- we call this dataset the \textbf{training dataset} -- and we feed it to our model. Each of the $i$-th data point was previously labelled, and in fact we feed each training data point $\mb{x}_i$ \emph{along} with its corresponding label $t_i$, so that the model can adapt its parameters accordingly to the \emph{target value} it is supposed to classify. This set of labels $\mb{t} = \left[ t_1 \; t_2 \; t_3 \; \dots \; t_N \right]$ is called the \textbf{target vector}.

@@

Furthermore, sometimes instead of applying the raw data to the model, we can apply some sort of preprocessing to the data to extract the relevant features from it. For instance, instead of just feeding a raw image, we can perform several operations like edge detection, and apply them in parallel.

The problems described above are, in fact, a subset of the problems that Machine Learning tries to address. These problems are called \textbf{classification problems}, because for each input data point, our model tries to fit it into the most appropriate class. But we can also address \textbf{regression problems} where the output is not limited to a discrete set of values but rather a continuous interval.

In summary, the most typical setting for a Machine Learning problem is having a large \textit{input dataset} which we use to \textit{train} our model (i.e. allowing him to dynamically adapt its set of parameters $\theta$), in order to produce an output label $y_i$ for each of them that minimizes a quality metric, typically defined as a \textit{cost function}. Artificial Neural Networks are one of the tools used in Machine Learning to perform this task, and will be discussed in Section~\ref{sec:theorBack_ann} as somewhat of a contextual introduction to the main theme of the thesis, which will be Recurrent Neural Networks (Section~\ref{sec:theorBack_rnn}), namely the \textbf{Long Short-Term Memory Networks} (Section~\ref{sec:theorBack_lstm}). These two last networks branch even further from these set of problems, and are usually employed in \textit{Deep Learning} tasks, where we try to extract even higher level information from data at the expense of increased model complexity.

@@

\begin{equation}\label{eq:update}
	w_{ji}^{(\tau+1)} = w_{ji}^{(\tau+1)} - \eta \frac{\partial E}{w_{ji}}
\end{equation}
we see that, from~\ref{eq:partialE}  it simply is the product of the error of the current neuron, $\delta_j$, with the output of the previous neuron $f(a_i)$. In turn, from~\ref{eq:deltaj}, we see the recurrence relationship between it and the weighted sum of all posterior neurons that connect to it. Hence, the name backpropagation is now clear: we are, in fact, propagating the errors backwards into the neuron of interest, weighted by the corresponding weight, but now \emph{backwards} instead of forward, as before. For the output units, the $\delta_j$ is simply the difference between the produced output and the corresponding label for that sample. This two-step process is performed for every data point in out dataset. %%Place where I can include an algorithmic overview?
For a complete proof of the above formulas, see~\citep[chap. 5.3.1]{Bishop2006}.

@@

Even though RNNs outperform static ANNs in sequence recognition problems~\citep{Bengio1991}, they fail to retain long-term dependencies. Of course that the weight training process is itself a form of memory, but the problem is that the weight update is much slower than the activations~\citep{Yoshua01}, and therefore this memory only retains short-term dependencies. This is because of the so-called \textbf{Vanishing Gradients Problem}~\citep{Yoshua94,Yoshua01}, where the error decays exponentially through time, and the impact of previous incoming data points on the training of the weights quickly decrease.

To overcome this issue, Hochreiter and Schmidhüber proposed, in 1997, a novel approach, called the Long-Short Term Network~\citep{Hoch97}. The next section explains the main structure of this approach, and also how it is trained, serving as a support basis for the work of this thesis.

>>>>

\section{Basic Concepts of Machine Learning}
Machine Learning is a field of Computer Science that studies the development of mathematical techniques that allow software to learn autonomously, without an explicit description of each rule of operation. Its goal is to extract latent features from the data that allow an immediate classification of each input data into a particular class -- the catch is that there is no previous rule formulation, but instead we have an adaptive model that adjusts is parameters accordingly with the input data it receives, improving the estimates it yields as it receives new input samples.

Let us consider a practical example. For instance, suppose we want to build a program that given an input audio waveform representation of a spoken word, it matches it into a particular word of a dictionary. We could, of course, devise a set of rules and exceptions for each word analysing some of its features (perhaps the Fourier representation of each one, and, from it, manually finding the appropriate rules for each), but apart from it being a very complex task, it wouldn't be a scalable solution, given the enormous number of words in each language. The approach taken by Machine Learning is diametrically different, and instead of manually processing each waveform, we build a large dataset, of size $N$, containing the waveforms of several words $\left[ \mb{x}_1 \; \mb{x}_2 \;\mb{x}_2 \; \dots \mb{x}_N \right]$ -- we call this dataset the \textbf{training dataset} -- and we feed it to our model. Each of the $i$-th data point was previously labelled, and in fact we feed each training data point $\mb{x}_i$ \emph{along} with its corresponding label $t_i$, so that the model can adapt its parameters accordingly to the \emph{target value} it is supposed to classify. This set of labels $\mb{t} = \left[ t_1 \; t_2 \; t_3 \; \dots \; t_N \right]$ is called the \textbf{target vector}.

@@

Sometimes, instead of applying the raw data to the model, we can apply some sort of \emph{preprocessing} to the data to extract the relevant features from it. For instance, instead of just feeding a raw image, we can perform several operations like edge detection or low-pass filtering, and apply them in parallel. In cases of highly dimensional data (i.e. each data vector has a very high number of features), we can apply techniques like \textbf{Principal Component Analysis} to reduce the feature space to a smaller dimension one, where the previous features were combined into two or three new features that pose themselves as the most relevant.

The problems described above are, in fact, a subset of the problems that Machine Learning tries to address. These problems are called \textbf{classification problems}, because for each input data point, our model tries to fit it into the most appropriate class. But we can also address \textbf{regression problems} where the output is not limited to a discrete set of values but rather a continuous interval. On the other hand, the Neural Network that this work will implement addresses a special kind of classification problem, where the classification decision is influenced not only by the current input sample, but also by a given \emph{window of samples} that trail the current sample.

In summary, the most typical setting for a Machine Learning problem is having a large \textit{input dataset} which we use to \textit{train} our model (i.e. allowing him to dynamically adapt its set of parameters $\theta$), in order to produce an output label $y_i$ for each of them that minimizes a quality metric, the \textit{Cost Funcion}, which can be chosen to the sum of squared differences, the log-loss, or any other appropriate  mathematical relationship  between the estimate $y_i$ and the correct label $t_i$.

Now that the basic Machine Learning concepts have been presented, I will discuss, henceforth, one of the most important algorithms that address the supervised classification problems, the \emph{Artificial Neural Networks} that will be discussed in Section~\ref{sec:theorBack_ann} as somewhat of a contextual introduction to the main theme of the thesis, which will be Recurrent Neural Networks (Section~\ref{sec:theorBack_rnn}), namely the \textbf{Long Short-Term Memory Networks} (Section~\ref{sec:theorBack_lstm}), both of which are improvements over the initial formulation of the ANNs. These two last networks branch even further from these set of problems, and are usually employed in \textit{Deep Learning} tasks, where we try to extract even higher level information from data at the expense of increased model complexity.

@@

\begin{equation}\label{eq:update}
	w_{ji}^{(\tau+1)} = w_{ji}^{(\tau)} - \eta \frac{\partial E}{\partial w_{ji}}
\end{equation}
we see that, from~\ref{eq:partialE}  it simply is the product of the error of the current neuron, $\delta_j$, with the output of the previous neuron $f(a_i)$. In turn, from~\ref{eq:deltaj}, we see the recurrence relationship between it and the weighted sum of all posterior neurons that connect to it. Hence, the name backpropagation is now clear: we are, in fact, propagating the errors backwards into the neuron of interest, weighted by the corresponding weight, but now \emph{backwards} instead of forward, as before. For the output units, the $\delta_j$ is simply the difference between the produced output and the corresponding label for that sample. This two-step process is performed for every data point in out dataset. %%Place where I can include an algorithmic overview?
For a complete proof of the above formulas, see~\cite[chap. 5.3.1]{Bishop2006}.

@@

Even though RNNs outperform static ANNs in sequence recognition problems~\cite{Bengio1991}, they fail to retain long-term dependencies. Of course that the weight training process is itself a form of memory, but the problem is that the weight update is much slower than the activations~\cite{Yoshua01}, and therefore this memory only retains short-term dependencies. This is because of the so-called \textbf{Vanishing Gradients Problem}~\cite{Yoshua94,Yoshua01}, where the error decays exponentially through time, and the impact of previous incoming data points on the training of the weights quickly decrease.

To overcome this issue, Hochreiter and Schmidhüber proposed, in 1997, a novel approach, called the Long-Short Term Network~\cite{Hoch97}. The next section explains the main structure of this approach, and also how it is trained, serving as a support basis for the work of this thesis.
